\section{Related Work}

\paragraph{Transformer inference and KV-cache memory.}
The Transformer architecture replaces recurrence with self-attention and is the basis of modern decoder-only LLMs \citep{vaswani2017attention}. During autoregressive decoding, systems cache previous key and value activations so that each new token can attend to the prefix without recomputing all prior layers. This cache turns attention from repeated full-prefix computation into an incremental memory problem: latency and throughput depend not only on floating-point work, but also on how many KV blocks must be stored, moved, paged, and read at each step. FlashAttention reduces attention memory traffic through IO-aware exact kernels \citep{dao2022flashattention}, while PagedAttention and vLLM manage KV memory in blocks to improve serving throughput and reduce fragmentation \citep{kwon2023efficient}. Other serving systems address batching, offloading, speculative execution, and distributed scheduling, including Orca, FlexGen, and SpecInfer \citep{yu2022orca,sheng2023flexgen,miao2023specinfer}. \uckv{} is complementary to these systems: it asks which cache entries should be retained under a reliability constraint, rather than only how retained entries should be stored or scheduled.

\paragraph{Token eviction and adaptive KV-cache retention.}
Many KV-compression methods rely on the observation that not all prior tokens remain equally useful. Scissorhands exploits the persistence of token importance during generation \citep{liu2023scissorhands}. H2O retains heavy-hitter and recent tokens and frames eviction through a dynamic submodular perspective \citep{zhang2023h2o}. StreamingLLM protects attention sinks and recent tokens to support stable long streaming inference \citep{xiao2024streamingllm}. SnapKV selects prompt tokens from observation-window attention before generation \citep{li2024snapkv}; PyramidKV allocates layer-wise budgets using a pyramidal information-funneling observation \citep{cai2024pyramidkv}; Ada-KV studies adaptive head-wise budget allocation \citep{feng2024adakv}. Recent methods further explore semantic clustering, retrieval heads, expected future attention, and benchmark-driven tradeoffs \citep{devoto2025expected,yuan2024kvbenchmark}. These methods supply strong retention primitives, but most choose a compression level from efficiency or average-quality objectives. \uckv{} instead treats compression as a selective memory action: aggressive eviction is allowed only when calibrated risk estimates indicate that the current context and generation state can tolerate it.

\paragraph{KV-cache quantization, streaming, and hybrid compression.}
A second line compresses activations rather than removing tokens. KIVI studies asymmetric low-bit KV quantization and observes different structure in keys and values \citep{liu2024kivi}. KVQuant pushes ultra-low-bit cache quantization for very long contexts with outlier-aware handling \citep{hooper2024kvquant}. CacheGen compresses and streams KV cache state across serving boundaries \citep{liu2024cachegen}. Hybrid approaches combine eviction, mixed precision, streaming, or layer-wise allocation, making the design space increasingly rich. This matters for \uckv{} because uncertainty calibration need not be tied to a single compression primitive: the risk gate can choose among token retention budgets, quantization levels, or conservative fallbacks. In the present draft we focus on token retention because it is easier to instrument in replay experiments, but the formulation is intended to cover mixed eviction--quantization policies.

\paragraph{Reliability risks of cache compression.}
The motivation for risk-aware compression is strengthened by recent evidence that cache decisions can distort long-context evaluation itself. Long-context benchmarks reveal that models can underuse evidence in the middle of a prompt or fail as context length increases \citep{liu2024lost,bai2024longbench,hsieh2024ruler,zhang2024infinitebench,li2024needlebench}. A compression method can therefore appear efficient on easy prompts while silently amplifying failures on position-sensitive retrieval, multi-hop reasoning, or rare evidence. Dedicated KV-compression benchmarks and risk analyses make this point explicit: compression quality must be evaluated not only by memory reduction, but also by what accuracy, robustness, and calibration are lost in return \citep{yuan2024kvbenchmark,haverbeck2026risk}. This draft follows that stance. We report small pilot results as preliminary and avoid claiming superiority until larger long-context and task-diverse evaluations are completed.

\paragraph{Uncertainty estimation, calibration, and risk control.}
Calibration asks whether predicted confidence matches empirical correctness probability. Temperature scaling is a standard post-hoc baseline for neural classifiers \citep{guo2017calibration}, and selective prediction studies the risk--coverage tradeoff when a model abstains under high uncertainty \citep{geifman2017selective}. In LLMs, confidence can come from token probabilities, verbalized confidence, self-evaluation, multiple samples, semantic entropy, or external detectors \citep{kadavath2022language,farquhar2024detecting,manakul2023selfcheckgpt,geng2024survey}. Conformal risk control provides a distribution-free way to set thresholds for bounded downstream loss under calibration data \citep{angelopoulos2024conformal}. These ideas are usually applied to final answers or abstention decisions. \uckv{} moves the same principle inside inference: the system abstains from aggressive memory compression when the calibrated risk of harming generation exceeds a target tolerance.

\paragraph{Positioning of \uckv{}.}
The closest recent direction to this draft is confidence-aware KV management, including work that uses next-token uncertainty to adapt cache budgets \citep{li2026confkv}. Our project differs in emphasis. First, it frames budget selection as a calibrated risk-control problem rather than a heuristic confidence trigger. Second, it separates three measurable quantities: memory savings, task degradation, and calibration error of the compression-risk predictor. Third, it keeps a conservative fallback path, so negative results are informative: if uncertainty gates do not dominate H2O-like baselines under matched budgets, the paper should report the boundary conditions rather than hide them. This makes the current manuscript a reliability-centered study of KV compression, not only an efficiency method paper.
